A free/open-source machine translation system for English to Kazakh

### 2015-03-04 engkaz-eamt-2015/engkaz-eamt-2015.tex
making changew
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59212 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{polyglossia}
\setdefaultlanguage[variant=australian]{english}
\usepackage{latexsym}
%\usepackage[small,bf]{caption}
\usepackage[bf]{caption}
\usepackage{xltxtra}
\usepackage{times}
\usepackage{fontspec}
\usepackage{footnote}
\usepackage{natbib}
\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
\usepackage{listings}

@@

\begin{abstract}
This paper presents rule-based machine translation from English to Kazakh. Main differences 
between two languages in syntax and morphology are presented, and showed how machine translation 
system was designed to solve these challenges. We showed an evaluation of system coverage and 
translation quality and future work.
\end{abstract}

@@

This paper presents a shallow-transfer rule-based machine translation system from English 
into Kazakh language. First of all, English language is a West Germanic language and Kazakh 
belong to group of Turkic languages. Therefore, the complex agglutinative morphology of Kazakh 
language is very different from morphologically not too complex language like English; on the other 
hand, Turkic language morphology shows clear morphotactics (ordering of morphemes), its morphophonology 
shows complex phonological changes to due to interactions between neighboring morphemes (vowel harmony, 
sonorisation, etc.) many of which are explicitly represented in writing.

On the other hand, there are many differences between the syntax of Turkic languages 
and English: subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of 
postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers 
always preceding the modified/specified (normally following in English), overt case marking allowing 
for a rather free ordering of arguments (versus a more fixed order in English), lack of definite 
articles (extensively used in English), verbal-noun-centered structures where English uses modal 
verbs (must, have to, want to) or verbal-noun or verbal-adjective-centred constructions where 
English has subordinate clauses using finite verbs with relatives or subordinating 
conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel 
of the English verb \emph{have}, as used for possession, etc. For an account (in Russian) of 
syntactic differences between English and Kazakh, see \cite{pecherskih2012}.

>>>>

\usepackage{polyglossia}
\setdefaultlanguage[variant=australian]{english}
\usepackage{latexsym}
%\usepackage[small,bf]{caption}
\usepackage[bf]{caption}
\usepackage{xltxtra}
\usepackage{times}
\usepackage{fontspec}
\usepackage{footnote}
\usepackage{natbib}
\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
\usepackage{listings}
\usepackage{microtype}

@@

\begin{abstract}
This paper presents the current state of development a shallow-transfer rule-based machine translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
between the two languages are presented, and it is showed how the MT 
system was designed to tackle these challenges. We show an evaluation of system coverage and 
translation quality and outline and future work.
\end{abstract}

@@

This paper presents a shallow-transfer rule-based machine translation (MT) system from English 
(a West Germanic language) to Kazakh (a Turkic language)
The complex agglutinative morphology of Kazakh 
is in contrast with the simple morphology of English.
Turkic language morphology shows clear morphotactics (ordering of morphemes), but with 
complex phonological changes (morphophonology) to due to morpheme contact (vowel harmony, 
sonorisation, etc.) many of which are explicitly represented in writing.

On the other hand, there are many differences between the syntax of Turkic languages 
and English: subject–object–verb order (vs.\ subject–verb–object in English), use of 
postpositions (vs.\ prepositions in English), head-final syntax with modifiers and specifiers 
always preceding the modified/specified (vs.\ following in English), overt case marking allowing 
for a rather free ordering of arguments (vs.\ a more fixed order in English), lack of definite 
articles (extensively used in English), verbal-noun-centered structures where English uses modal 
verbs (\emph{must}, \emph{have to}, \emph{want to}) or verbal-noun or verbal-adjective-centred constructions where 
English has subordinate clauses using finite verbs with relatives or subordinating 
conjunctions (\emph{the book which I read}, \emph{the place where I saw him}, \emph{before he came}), lack of a parallel 
of the English lexical verb \emph{to have}, etc. For an account (in Russian) of 
syntactic differences between English and Kazakh, see \cite{pecherskih2012}.
